---
title: "Practical ML"
author: "SENG KRUY"
date: "6/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data Processing
First of all, we load training and testing data from files. From the data, we remove unrelated features such as id number, name, timestamp etc. We also remove feature which consist of null record. We did the same to test set.
```{r cars}
library(caret)
train_set<-read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!"))
test_set<-read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!"))

# Remove unrelated features

NA_Count = sapply(1:dim(train_set)[2],function(x)sum(is.na(train_set[,x])))
NA_list = which(NA_Count>0)
train_set = train_set[,-NA_list]

train_set = train_set[,-c(1:7)]
test_set = test_set[,-c(1:7)]
train_set$classe = factor(train_set$classe)

```
We convert the label of training set to categories. Below is the number of row belong to each label:
``` {r }
plot(train_set$classe)
```
## Model Training
We choose to train our model with a famous Gradient Boosting Machine (GBM). Some advantages of GBM are:

1. An extremely popular machine learning algorithm that have proven successful across many domains and is one of the leading methods for winning Kaggle competitions. 
2. Lots of flexibility - can optimize on different loss functions and provides several hyperparameter tuning options that make the function fit very flexible.
3. No data pre-processing required - often works great with categorical and numerical values as is.
4. Handles missing data - imputation not required.

We train GBM model with 4 fold cross validation (to save time). Grid search on some parameters such as depth and n.tree is also applied to fine tune for optimal parameter. 
```{r , include=FALSE}
# Cross Validation with Random Forest 

cvctr = trainControl(method="cv",number=3,allowParallel=TRUE,verboseIter=TRUE)

gbmGrid <-  expand.grid(interaction.depth = c(1,5,10,20,30), 
                        n.trees = c(10,50,100,150,200), 
                        shrinkage = 0.1,
                        n.minobsinnode = 20)
gbm_model <- readRDS("gbm_model.rds")
#gbm_model = train(classe~., data=train_set, method="gbm",trControl=cvctr,tuneGrid=gbmGrid)
```
Below is plot showing the accuracy of our model with different depth and n.tree. The result shows that the model can acheive the highest accuracy at the max tree depth of 30. Similar accuracy results are also found for model with max tree depth of 20 and 10 at 200 iteration (n.tree).

```{r pressure, echo=FALSE}
plot(gbm_model)
```

Below is the confusion matrix tested on training set. We found that the model is able to predict correctly on all classes.

```{r , echo=TRUE}
prf=predict(gbm_model,train_set)
table(prf,train_set$classe)

```

## Test on testing set
```{r , echo=TRUE}
predictions <- predict(gbm_model,newdata=test_set)
predictions
```